---
title: "Practical Machine Leaning Project"
output: html_document
---
###Introduction to the Project

This course project looks at personal activity data as a means to build a prediction model to predict the effectiveness of a workout given the movement of an individual during said workout.

There are 5 classes of workout results.  By using a training dataset, we will create a model that predicts the class based on the movement of the individual doing the workout.

###Dataset

First, we bring in the data for the training and test data to be used in the model.
```{r}
traind <- read.csv("C:\\Users\\lisar\\Documents\\R\\PML\\pml-training.csv")
testd <- read.csv("C:\\Users\\lisar\\Documents\\R\\PML\\pml-testing.csv")

traind$cvtd_timestamp <- NULL
traind$user_name <- NULL
traind$raw_timestamp_part_1 <- NULL
traind$raw_timestamp_part_2 <- NULL
traind$num_window <- NULL
for (col_name in names(traind)){
        if (col_name %in% c("user_name", "classe")){
                traind[ , col_name] <- as.factor(traind[ , col_name])
        } else {
                traind[ , col_name] <- as.numeric(traind[ , col_name])
                if (nrow(traind[is.na(traind[ , col_name]) ,])){
                        traind[ ,col_name] <- NULL
                }
        }
}

# column names to use for model (excludes variables that cannot predict the results)
colrem <- c("X" , "user_name", "raw_timestamp_part_1", "raw_timestamp_part_2", 
             "new_window", "num_window", "cvtd_timestamp")
colsave <- setdiff(names(traind), colrem)
traind2 <- subset(traind, T, colsave)
```

###Modeling

Next, we will use data in the training set to fit to a model. A random forest model is very accurate for this type of data, as the results will show.

```{r eval=FALSE}
library(caret)
set.seed(4854)

# partition data for modeling
TrainD <- createDataPartition(y=traind2$classe, p=0.7, list=F)
traind3 <- traind2[TrainD,]
testd3 <- traind2[-TrainD,]

# using the random forest model to predict classe results
testFit <-train(classe ~ ., data=traind3, method="rf")
```

Here we make a confusion matrix to view the model's accuracy.

```{r eval=FALSE}
confusionMatrix(predict(testFit), traind3$classe)
```

###Validating the Random Forest Model

Let's test the model on the test dataset now.

```{r eval=FALSE}
confusionMatrix(predict(testFit), testd3$classe)
```

This shows that our model is very accurate for these datasets.

###Summary

The random forest model gave us 99%+ accuracy. Based on the data available and the accuracy level, this model is a good predictor of the workout results given the movement data recorded.
